### 2013-12-12 Goal/goals.tex
Initial commit

\documentclass{book}
\usepackage[utf8]{inputenc}
\usepackage[dutch]{babel}
\usepackage{hyperref}

\begin{document}
\chapter{Probleem- en doelstelling}
\section{Probleemstelling}

\section{Doelstelling}
\end{document}