### 2014-02-01 Thesis/hoofdstukken/uitrol.tex
Thesistekst begin

\chapter{Automatisch uitrollen van databases}
\section{Inleiding}
%Inleiding

\section{IMP: Framework voor installatie}
%Uitleg van IMP

%Uitleg van hoe voor elk systeem

%