### 2014-07-05 Thesis - 2/overige/afkortingen.tex
Schrijven van geschiedenis

\chapter{Lijst van afkortingen en symbolen}
\section*{Afkortingen}
\begin{flushleft}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabularx}{\textwidth}{@{}p{12mm}X@{}}
    IMP   & Infrastructure Management Platform \\
    DBMS   & Databasemanagementsysteem \\
  \end{tabularx}
\end{flushleft}
\section*{Symbolen}
\begin{flushleft}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabularx}{\textwidth}{@{}p{12mm}X@{}}
    42    & aaa \\
  \end{tabularx}
\end{flushleft}