### 2014-07-22 Thesis - 2/overige/afkortingen.tex

\chapter{Lijst van afkortingen en symbolen}
\section*{Afkortingen}
\begin{flushleft}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabularx}{\textwidth}{@{}p{12mm}X@{}}
    ACID & Atomicity, consistency, integrity en durable.  \\
   	BASE & Basic Availability, Soft state en eventuel consistency.  \\
    DBMS   & Database management systeem \\
    RDBMS   & Relationeel Database management systeem \\
    IMP   & Infrastructure Management Platform \\
    YCSB & Yahou Cloud Serving Benchmark \\

  \end{tabularx}
\end{flushleft}